---
title: "R Exercises"
date: "October 27, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
gapminder <- read.csv("https://raw.githubusercontent.com/swcarpentry/r-novice-gapminder/gh-pages/_episodes_rmd/data/gapminder_data.csv", stringsAsFactors = FALSE)
meanLifeExp <- mean(gapminder$lifeExp)


```

# Exercise 1
Q1. Which of the following will NOT return TRUE?

>A. FALSE == FALSE  
B. 10-5 == sqrt(25)  
C. TRUE > FALSE  
D. 'a' > 'b'

Q2. What is the output when we execute the following code?
x <- 3
y <- 2
y <- 17.4
x+y

>A. [1] 3  2  17.4  
B. [1] 22.4    
C. [1] 20.4   
D. [1] 5  

\pagebreak

# Exercise 2
Exercise:
Import the gapminder data frame again.
Use `str()` to look at the structure of the dataframe and `summary()` to get information about the variables.

>What are its columns?  
How many rows and columns are there?  
What is the earliest year in the `year` column?  
What is the average life expectancy?  
What is the largest population?  


\pagebreak


# Exercise 3
### 3a. How would we look at life expectancy overtime with ggplot?
```{r}
ggplot(data = gapminder, mapping = aes(x = year, y = lifeExp)) +
  geom_point()
```

\pagebreak
### 3b. Try to add color.  
There is a color argument for the aes function  
You can can tell ggplot to color by continent  

```{r}
ggplot(data = gapminder, mapping = aes(x = year, y = lifeExp, color = continent)) +
  geom_point()
```

\pagebreak
### 3c. Switch the order of the geom_ functions. What happened?
```{r}
ggplot(data = gapminder, mapping = aes(x=year, y=lifeExp, by=country)) +
  geom_point() + geom_line(mapping = aes(color=continent))
```

\pagebreak
### 3d. Change the color and size of the points outside of aes.   
Note: ?geom_point

```{r}
ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp)) +
  geom_point(size=3, color="green") + scale_x_log10() +
  geom_smooth(method="lm", size=1.5)
```


\pagebreak
### 3e. Now change the shape of the points and color the points by the continent  
Note: You'll get multiple linear model fits  
Note: ?geom_point  
```{r}
ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp, color = continent)) +
  geom_point(size=3, shape=17) + scale_x_log10() +
  geom_smooth(method="lm", size=1.5)
```  

\pagebreak
### 3f. Create a density plot of GDP per capita, filled by continent.  
Advanced: Transform the x axis to better visualise the data spread.  
Add a facet layer to panel the density plots by year.  

Answer:
```{r}
ggplot(data = gapminder, mapping = aes(x = gdpPercap, fill=continent)) +
  geom_density(alpha=0.6)# + facet_wrap( ~ year) + scale_x_log10()
```  
Advanced answer:
```{r}
ggplot(data = gapminder, mapping = aes(x = gdpPercap, fill=continent)) +
  geom_density(alpha=0.6) + facet_wrap( ~ year) + scale_x_log10()
```  


# Exercise 4: loops
### Write a script that loops through the gapminder data by continent and prints out whether the mean life expectancy is smaller or larger than 50 years.

Answer
```{r}
meanLifeExp <- mean(gapminder$lifeExp)

for(cont in unique(gapminder$continent)){
  if(mean(gapminder[gapminder$continent == cont, 'lifeExp']) > meanLifeExp){
    print(paste(cont, 'lifeExp is above average'))
  }
  else if (mean(gapminder[gapminder$continent == cont, 'lifeExp']) < meanLifeExp){
    print(paste(cont, 'lifeExp if below average'))
  }
  else{
    print(paste(cont, 'lifeExp is average'))
  }
}
```

